---
title: "Fix MP Names"
author: "Matthew Zatzman"
date: "`r Sys.Date()`"
format: 
  html:
    code-fold: true
    code-tools: true
    code-link: true
    toc: true
---


```{r}
library(tidyverse)
library(HGNChelper)
```

The Tirosh metaprograms from https://www.nature.com/articles/s41586-023-06130-4 contain several unapproved HGNC names. Here we catalog and fix them.

```{r}
#| warning: false
#| message: false
meta_prog_file <- "~/work/references/meta_programs_tirosh/meta_programs.xlsx"

meta_progs_fix <- map(excel_sheets(meta_prog_file), .f = function(sheet) {
  df <- read_xlsx(path = meta_prog_file, sheet = sheet) %>%
    janitor::clean_names() %>%
    as.list()
    
  
  res <- map(names(df), .f = function(x) {
    fix <- HGNChelper::checkGeneSymbols(df[[x]])
    fix$ct <- sheet
    fix$mp <- x
    colnames(fix)[1] <- "Gene"
    fix$gene_rank <- 1:nrow(fix)
    return(fix)
  }) %>%
    list_rbind()
  return(res)
}) %>%
  list_rbind()
```

```{r}
# Proportion of genes that do not map
sum(!meta_progs_fix$Approved) / nrow(meta_progs_fix)
```

List all genes that have unapproved symbols

```{r}
meta_progs_fix %>%
  filter(!Approved) %>%
  select(Gene, Approved, Suggested.Symbol) %>%
  distinct() %>%
  DT::datatable()
```

Try patching with Seurat function (nothing gets patched)

```{r}
non_mapped <- filter(meta_progs_fix, is.na(Suggested.Symbol))$Gene %>% unique()

seurat_try <- Seurat::UpdateSymbolList(non_mapped)
```


Write out table to manually patch where possible

```{r}
meta_progs_fix %>%
  filter(!Approved) %>%
  select(Gene, Approved, Suggested.Symbol) %>%
  distinct() %>%
  arrange(Suggested.Symbol, Gene) %>%
  write.table("non_mapping_genes_unedited.txt", quote = F, sep = "\t", col.names = T, row.names = F)
```

Outside this script I manually select where there are multiple potential matches. One manual patch was in `ZFP385A` to `ZNF385A` Now reload and format

```{r}
patched_names <- read.table("non_mapping_genes_edited.txt", header = T, sep = "\t") %>%
  filter(!is.na(Suggested.Symbol))

# Patch in
meta_progs_fix[match(patched_names$Gene, meta_progs_fix$Gene), "Suggested.Symbol"] <- patched_names$Suggested.Symbol

# How well do we do
sum(is.na(meta_progs_fix$Suggested.Symbol)) / nrow(meta_progs_fix)
```

Save out flat table

```{r}
meta_progs_fix %>%
  dplyr::rename("original_symbol" = "Gene",
         "Gene" = "Suggested.Symbol",
         "cell_type" = "ct",
         "meta_program" = "mp") %>%
  relocate(Gene) %>%
  write.table("tirosh_mp_patched.txt", quote = F, sep = "\t", row.names = F, col.names = T)
  
```




